How to Analyze Build Logs---A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-03-13 tse-paper/abstract.tex
More corrections

Software builds from continuous integration produce detailed logs about the status and results of the various tools involved in the build.
These build logs are a valuable data source for developers and researchers to inspect test results, the duration of build steps or understand the cause of a build failure.
However, build logs are very verbose, at best semi-structured and their structure differs highly between projects.
This makes it hard to process and analyze them.
In this paper, we evaluate and compare three different techniques that aim to retrieve specified text chunks from a build log, namely program synthesis by example, textual similarity and search keywords.
We conduct an empirical study by comparing these techniques on the \emph{LogChunks} data set of 797 Travis CI logs from a diverse range of projects.
Our findings show that none of the three techniques clearly outperforms the others.
We discuss under which circumstances each technique performs best and provide a recommendation on when developers or researchers should use which technique.

>>>>

Build logs are a textual by-product that software build processes
create, often as part of their Continuous Integration (CI)
pipeline. Build logs can serve as a paramount source of
information for developers to understand and debugg build or
test failures. Manually
extracting the important chunks of information, though, is akin to
finding a needle in a haystack. Recently, researchers and practitioners have begun attempts to partly automate
this time-consuming activity, with the proposition to not only ease developers' tasks but to enable a new class of automated on-ward processing applications.
In this paper, we want to give a systematic overview of the state of the art of the emerging field of build log analysis. To this end, we first survey and categorize the existing methods to extract information chunks from build logs. We then develop prototypical implementations for three promising techniques, namely program synthesis
by example, textual similarity and search keywords. Finally, we evaluate the three techniques in an empirical study on our manually
labeled \emph{LogChunks} data set, which comprises 797 build logs in 29 languages.
Our findings show that none of the three techniques in general outperforms
the others, but rather, that each technique has its respective strengths and weaknesses. We discuss under which circumstances each technique performs best
and provide a recommendation on when developers or researchers should use which
technique.

### 2020-03-13 tse-paper/content.tex
More corrections

The goal of this paper is to investigate different chunk retrieval techniques for build logs and describe under which circumstances certain techniques can be recommended over others.
We aim to characterize different chunk retrieval techniques.
For \textbf{Research Question 1} (\textbf{RQ1}), we analyze which criteria influence the suitability of a chunk retrieval technique for CI build logs.

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item program synthesis by example using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as CTS), and
  \item keyword search (referred to as KWS).
\end{itemize}
\textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
\textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

@@

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision and
recall of the chunk retrieval is required.
CTS is well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by a human.
In cases where the textual representation of the desired information in the log
is unpredictable or varies greatly, KWS is the best technique to choose.
However, its low precision requires a human to interpret the output of the chunk
retrieval.

@@

\section{Related Work}
\label{sec:rw}
We explain how researchers gather information about CI usage in software projects
through build log analysis and why the chunk retrieval techniques we investigate
simplifies their data collection.
This section moves on to past works about augmenting build logs to make it easier
for developers to inspect them.
The presented approaches are related to and can benefit from the techniques we
analyze in this paper.
Further, we classify build logs as semi-structured data and differentiate our
work from system log analysis.

\subsection{Build Failures in CI}
Various researchers look into why CI builds fail and into the impact of build failures on the development workflow.

@@

\noindent
\textbf{Supporting Log Analysis with Chunk Retrieval}
To leverage the valuable information within build logs the researchers presented
in this section create parsers and regular expression-based programs.
This task of retrieving specific chunks of text from the build logs can be
solved by the chunk retrieval techniques we compare in this paper.
Our results can support researchers in choosing a suitable technique for their
data set of build logs and the chunks they want to retrieve.
By relieving them from building custom parsers we enable them to cover a much
wider range of languages and build tools in their studies.

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why their build failed.
Several researchers are looking into supporting developers to process the verbose build logs.
Vassallo et al.~\cite{vassallo2018un-break} try to shorten the time it takes developers to understand build logs.
They parse Maven~\cite{maven2019website} build logs into a structured representation and create hint generators.
%The hint generators leverage this structured access to the information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for discussions related to why the build failed.
In a qualitative study they observed that highlighting the locality and context of an issue is helpful to programmers.
%Their tool BART is published as a Jenkins Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this paper can be used to fill similar structured representations with information from build logs.
As they simplify the construction of parsers they would enable developers and researchers to cover a wider array of build tools, which is the main influence factor on the structure of a build log.

@@

The first goal in parsing system log files is to separate constant and variable parts within a log message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log messages are clustered into log events, unifying messages with identical constant parts and varying parameters.
The output of a log parser is a structured log, composed of a list of timed events and the corresponding parameter values~\cite{he2016evaluation}.
This structured log is then the input to various machine learning and data mining processes.
%Researchers mine patterns for operational profiling~\cite{nagappan2009efficiently}, debugging~\cite{oliner2012advances}, performance analytics or anomaly detection~\cite{nagappan2010abstracting}.
%Xu et al.~\cite{xu2009detecting} leverage the connection of log statements to the source code producing them to separate messages into constant and variable parts more accurately.

The techniques developed for system log analysis can also be applied to build logs.
One example is comparing execution traces to reference traces of intended behavior to detect anomalies.
Amar et al.~\cite{amar2019mining} employed a similar approach to detect relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of events.
%A similar approach could also be applied to build logs to determine the sequence of executed build steps or phases.
In this paper we focus on extracting a single specified information from the build log as a whole with chunk retrieval techniques.
Chunk retrieval techniques are used as a part of log parsing to retrieve the values of variable parts in a log message, e.g.\ by using regular expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
	basicstyle=\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet- Responder: Received block blk_3587508140051953248 of size 67108864 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140051953248 terminating
  \end{lstlisting}  
  \caption{System Log Statements. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure}

@@

Programming by Example enables end users to automate repetitive tasks.
The user provides examples for the input and the corresponding output and a
synthesis algorithm tries to create the program intended by the user.
This section introduces the theoretical foundations of the program synthesis
algorithm of the \emph{PROgram Synthesis using Examples} (PROSE) framework~\cite{prose2019webpage}.
The PROSE framework is developed by Microsoft Research.
Next, this section presents the FlashExtract DSL, which defines text extraction
tasks within PROSE and is the basis for the implementation of our chunk retrieval technique PBE\@.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces the concept of chunk retrieval techniques.
We use these techniques to extract text chunks from build logs that represent a specific, targeted information.
We present the three chunk retrieval techniques we investigate in this paper:
program synthesis by example (PBE), common text similarity (CTS) and keyword search (KWS).

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
For this paper, we want to evaluate different techniques to retrieve information chunks from build logs, which we call \emph{chunk retrieval techniques}.
The techniques we investigate do not require to parse the structure of a whole build log, but focus on extracting just one specific information.

The user provides a \textit{configuration} that specifies which information the chunk retrieval should target and supplies the necessary information for the technique to identify the targeted information chunk in a build log.
Each chunk retrieval has a specific \textit{granularity}, i.e.\ the smallest retrievable text piece and uses a specific \textit{identification technique} to select the log parts it retrieves.
Each configuration addresses a specific \textit{scope}.
The scope can be a specific project, a tool involved in the build, a programming language or global, when configuring a retrieval technique for all possible build logs.
A \emph{run} of a chunk retrieval technique consumes a build log as a plain text file and produces a string output, which consists of substrings of the build log text.

@@

\noindent
\textbf{Training and Test Set}
We use \emph{LogChunks} as the data set for our study.
For each one of 80 repositories it contains about 10 build logs, manually labeled with the substring describing the reason the build failed.
In addition to that, it contains keywords to search for that substring and which structural category the substring belongs to.

For each repository in \emph{LogChunks}, we split the examples chronologically into training and test set.
Therefore, we train on examples from past build logs and test on more recent build logs.

>>>>

The goal of this article is to investigate different chunk retrieval techniques for build logs and describe under which circumstances certain techniques can be recommended over others.
We aim to characterize different chunk retrieval techniques.
For \textbf{Research Question 1} (\textbf{RQ1}), we analyze which criteria influence the suitability of a chunk retrieval technique for CI build logs.

We implement and evaluate three chunk retrieval techniques:
\begin{itemize}
  \item program synthesis by example using the Microsoft PROSE library (referred to as PBE),
  \item a common text similarity approach (referred to as CTS), and
  \item keyword search (referred to as KWS).
\end{itemize}
% TODO: Shortly describe each of these techniques

\textbf{RQ2} asks under which conditions PBE, CTS and KWS are suited to retrieve information from continuous integration build logs.
\textbf{RQ2} is refined into sub-questions along with the criteria resulting from \textbf{RQ1} and compares their instantiations for the three techniques:
how many training examples a technique needs to perform best (\textbf{RQ2.1}), how structurally diverse the examples can be (\textbf{RQ2.2}) and how accurate the retrieved output is (\textbf{RQ2.3}).
To evaluate PBE, CTS and KWS we use the \emph{LogChunks} data set, which encompasses about 800 log files from 80 repositories.
Each log is labeled with the log part describing the reason a build failed, keywords to search for this log part and a categorization of the labeled log part according to its structural representation within the log.

% TODO: Add structural category explanation
% TODO: Add running example

@@

We recommend PBE for use cases where the desired information is always
represented in the same structural way and high confidence in precision and
recall of the chunk retrieval is required.
CTS is well suited when the representation of the desired information varies
slightly and the output of the chunk retrieval is further processed by a human.
In cases where the textual representation of the desired information in the log
is unpredictable or varies greatly, KWS ssems to be the best choice. However, its low precision---it extracts a context of multiple lines around a find---makes it generally unsuited for automatic on-ward processing and instead requires a human to interpret the output chunk.

@@

\section{Related Work}
\label{sec:rw}
We explain how researchers gather information about CI usage in software projects
through build log analysis and why the chunk retrieval techniques we investigate
simplifies their data collection.
This section moves on to past works about augmenting build logs to make it easier
for developers to inspect them.
The presented approaches are related to and can benefit from the techniques we
analyze in this article.
Further, we classify build logs as semi-structured data and differentiate our
work from system log analysis.

\subsection{Build Failures in CI}
Various researchers studied why CI builds fail and into the impact of build failures on the development workflow.

@@

\noindent
\textbf{Supporting Log Analysis with Chunk Retrieval}
To leverage the valuable information within build logs the researchers presented
in this section create parsers and regular expression-based programs.
This task of retrieving specific chunks of text from the build logs can be
solved by the chunk retrieval techniques we compare in this article.
Our results can support researchers in choosing a suitable technique for their
data set of build logs and the chunks they want to retrieve.
By relieving them from building custom parsers we enable them to cover a much
wider range of languages and build tools in their studies.

\subsection{Augmentation of Build Logs}
\label{sec:rw-bl-analysis}
Build logs are a valuable data source for developers to find out why their build failed.
Several researchers are looking into supporting developers to process the verbose build logs.
Vassallo et al.~\cite{vassallo2018un-break} try to shorten the time it takes developers to understand build logs.
They parse Maven~\cite{maven2019website} build logs into a structured representation and create hint generators.
%The hint generators leverage this structured access to the information within the build log to propose fixes.
For example, one of the hint generators queries stack overflow for discussions related to why the build failed.
In a qualitative study they observed that highlighting the locality and context of an issue is helpful to programmers.
%Their tool BART is published as a Jenkins Plugin~\cite{bart2019plugin}.
The chunk retrieval techniques we compare in this article can be used to fill similar structured representations with information from build logs.
As they simplify the construction of parsers they would enable developers and researchers to cover a wider array of build tools, which is the main influence factor on the structure of a build log.

@@

The first goal in analyzing system log files is ubiquitously to separate constant and variable parts within a log message~\cite{nagappan2010abstracting,he2017towards}.
Next, the log messages are clustered into log events, unifying messages with identical constant parts and varying parameters.
The output of a log parser is an ordered list of timed events and their corresponding parameter values~\cite{he2016evaluation}.
This structured log is then the input to various machine learning and data mining processes.
%Researchers mine patterns for operational profiling~\cite{nagappan2009efficiently}, debugging~\cite{oliner2012advances}, performance analytics or anomaly detection~\cite{nagappan2010abstracting}.
%Xu et al.~\cite{xu2009detecting} leverage the connection of log statements to the source code producing them to separate messages into constant and variable parts more accurately.

The techniques developed for system log analysis can also be applied to build logs.
One example is comparing execution traces to reference traces of intended behavior to detect anomalies.
Amar et al.~\cite{amar2019mining} employed a similar approach to detect relevant lines in build logs.
%Classic log parsers interpret the whole log file into a sequence of events.
%A similar approach could also be applied to build logs to determine the sequence of executed build steps or phases.
In this article, we focus on extracting a single specified information from the build log as a whole with chunk retrieval techniques.
Chunk retrieval techniques are used as a part of log parsing to retrieve the values of variable parts in a log message, e.g. by using regular expressions~\cite{nagappan2010abstracting,xu2009detecting}.

\lstset{
  morekeywords={INFO, WARN, 2008-11-09},
	basicstyle=\scriptsize,
  postbreak=\mbox{\textcolor{cyan}{$\hookrightarrow$}\space},
  showspaces=false,
  showstringspaces=false,
  keywordstyle=\color{blue},
	frame=single,
  extendedchars=false,
  texcl=false
}
\begin{figure*}[!t]
  \centering
  \begin{lstlisting}[breaklines=true]
2008-11-09 20:46:55,556 INFO dfs.DataNode$Packet- Responder: Received block blk_3587508140051953248 of size 67108864 from /10.251.42.84
2008-11-09 20:49:46,764 WARN PacketResponder 0 for block blk\_3587508140051953248 terminating
  \end{lstlisting}  
  \caption{System Log excerpt. Example adapted from~\cite{he2017towards}.}
  \label{lst:system-log}
\end{figure*}

@@

Programming by Example (PBE) is a form of generalized program synthesis that enables users to automate repetitive tasks, without them having to explicitly code a program. The users provide examples for the input and the desired output and a
synthesis algorithm tries to create a program that transforms the input to the output, hopefully as close as possbile to the user's intention.
This section introduces the theoretical foundations of the program synthesis
algorithm of the \emph{PROgram Synthesis using Examples} (PROSE) framework~\cite{prose2019webpage}.
The PROSE framework is developed by Microsoft Research.
Next, this section presents the FlashExtract DSL, which defines text extraction
tasks within PROSE and is the basis for the implementation of our chunk retrieval technique PBE\@.

@@

\section{Chunk Retrieval Techniques}
\label{sec:techniques}
This section introduces key concepts of the three chunk retrieval techniques we study in this article, namely program synthesis by example (PBE), common text similarity (CTS), and keyword search (KWS).
We use these techniques to extract text chunks from build logs that represent a specific, targeted information.

@@

\subsection{Characteristics of Chunk Retrieval Techniques}
\label{sec:blirt}
In this article, we evaluate different techniques to automatically retrieve pieces of information that appear literally in the build logs, i.e., techniques that do not aggreagte, combine, or deducted information. We call such pieces of information from build logs \emph{chunks}, and the techniques \emph{chunk retrieval techniques}.
The techniques we investigate here do not require to parse the entire structure of build logs, but focus on extracting just one specific piece of information per configuration.

With the term \textit{configuration,} we want to abstract over the training and parametrization that techniques require in different forms.
The user provides a \textit{configuration} that specifies which information the chunk retrieval should target and supplies the necessary information for the technique to identify the targeted information chunk in a build log.
Each chunk retrieval has a specific \textit{granularity}, i.e.,\ the smallest retrievable text piece and uses a specific \textit{identification technique} to select the log parts it retrieves. The granularity might be adjustable by configuration.
A \emph{run} of a chunk retrieval technique consumes a build log as a plain text file and produces a string output, which consists of substrings of the build log text.

@@

\noindent
\textbf{Training and Test Set}
We use \emph{LogChunks} as the data set for our study.
For each of the 80 repositories in it, it contains 10 build logs, manually labeled with the substring describing the reason the build failed.
Additionally, it contains keywords to search for that substring and which structural category the substring belongs to.
% TODO (Caro): Add details how key words were extracted

% TODO (Caro): Add more details. How was the split in numbers?
For each repository in \emph{LogChunks}, we split the examples chronologically into training and test set.
Therefore, we train on examples from past build logs and test on more recent build logs.
